Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach

### 2017-05-21 answer_to_second_referee_comments.tex

We thank the reviewer for his/her helpful comments and apologize for the delay in returning a revised version of the manuscript. Changed text in the manuscript appears in bold. Our responses to individual comments are in the text below, beginning with ---

%TBD: make sure to delete comments in this file before submitting!

@@

Reviewer's Comments:

This paper proposes the use of self-organizing maps (1D and 2D) to aid in classifying galaxy spectra.  The revision has addressed several points and improve the clarity of the text.  However, there are several remaining issues.

1. The authors did not address my comment about the significance of the work.  What was learned?  What method is advocated, and what is the evidence for its superiority?

The sentence that was added about 1D vs. 2D maps comes very late in the paper and is quite abstract, stating that the two methods are complementary.  The Introduction should establish the goals of the investigation so the reader understands the purpose of each experiment as well as its significance.

--- We have reorganized the last two paragraphs of Section 1 to better describe the goals of the investigation. 
%TBD: 
We have also added some text to Section 5 discussing the significance.

The paper claims that "Compared to previous attempts to classify this sample, the self-organizing map approach was able to better classify the spectra which had similarities with more than one template." (abstract), but I cannot find a quantitative comparison with any previous results.  Figure 9 is presented as the result to compare with previous work, but since the previous results are not presented nor summarized, the reader can't tell how they differ.  The paper reports "more precision", but this is not defined; does it mean smaller error bars?  Stronger correlation?  How does this show "better classification"?  To support such a claim, this paper needs to select a performance metric and report on the values achieved in these experiments, compared to other methods.
%--- should I put the figure from the previous paper, maybe Noll09 one??! 
% I don't think including the figure from Noll09 will be enough. It's the performance metric that's critical here.

I had suggested comparing the results to a simple nearest-neighbor classification baseline, but this was not addressed.
%--- why could not I find anything about it in the 1st report?:(
%20170122PB: it's there but well-hidden - the sentence which includes "what performance does one get simply by assigning each of the 142 spectra to its closest match from the 12 templates"
--- We apologize for missing this suggestion in the previous version. We have now included this in the analysis and the results are reported in Section XX.

The comment about how SOMs provide an ordered clustering of the data (with respect to feature values), which k-means does not, is certainly true.  But the paper must then explain why this is important or valuable.
--- We have added text to the end of both Section 4.1 and Section 5 explaining the utility of an ordered clustering.

2. Perhaps I have missed it somewhere, but I don't think the paper ever states how exactly the classification of new data is done.  As I understand it, the authors train an SOM using template spectra, then classify new spectra by determining which SOM node is the best fit for the new item and classifying it based on the labeled template spectra that also reside in that node -- but this isn't actually stated anywhere.  Please add text explaining the classification step.  This would probably go in the final paragraph of the introduction and/or a new subsection in section 3 ("Method") - this section describes SOMs and k-means but not how to use either model to classify new data. (i.e., the SOM can't classify anything without knowing the labels of its original data set.  So although it builds the SOM in an unsupervised way, it requires the supervision of the template labels to be able to assign a label to a new item).
-- Added the following paragraph  at the end of section 3.2: 
    "As mentioned in the introduction, one of the main advantages of the SOM method  is that the resulting networks can be used to cluster new datasets with no additional training required. 
     New data with the same dimensionality as the original input data can be presented to the already trained network: the SOM algorithm finds the best matching unit considering the weight of nodes in the trained network and each vector from the new data set.
     The winner node determines the place of each new vector on the SOM.
     The location of the new vectors on the SOM allows them to be compared with the original data set. 
     Sections~\ref{sec: 1Dv} and ~\ref{sec: 2D} show the results of galaxy spectral classifications using networks trained on template spectra."

On page 2, lines 29-31, the authors added this text: "An SOM can be trained without supervision and its results used to classify other data. Therefore, this method can be considered as (semi)-supervised." As written, the phrase "this method" seems to refer to SOMs (so the statement wouldn't be true - SOMs are fully unsupervised), but I think instead the authors' proposed method (of building an SOM and then applying it to classify new data, using labels on the SOM-grouped data) is what is meant.  Please clarify.
--- Changed to: 
    "The training of a self-organizing map is fully unsupervised.
     Using the resulting map as a template to classify other data requires labelling the output, which is why some groups consider self-organizing maps to be a (semi)-supervised method."

3. The inclusion of k-means helps give useful context to this work. Figure 10 shows a head-to-head comparison of k-means and SOM applied to the 12 templates using 4 clusters/nodes.  But I can't tell which one is better.  The text states that they are similar but "show discrepancies for spectra types between the extreme cases".  Which ones?  How can I tell from this figure?
--- % need a comment here

At the bottom of page 10, the paper states: "The other difference between these two methods is that the network created by the SOM method can be used to classify the other spectra, while for the K-means clusters they can only be used for one set of data."  This isn't true.  (The paper itself points out that SOM reduces to k-means if the weight on the neighborhood function goes to 0, so you should be able to use the result in the same way.)  You can (and should, to do a direct comparison) use the clusters just like you used the SOM nodes - for a new data point, assign it to the most similar cluster center, then assign the template label from the original data used to construct the clusters.
%--- I do not know how to do that!
% for each spectrum to be classified, compute the distance (in whatever metric we end up using)
% between the spectrum and the K cluster centres. Spectrum is classified as belonging to whatever cluster results in the smallest distance.

To analyze the 142 galaxies, the SOM and k-means experiments used different methodology and so are difficult to compare or understand. It seems that for k-means, the authors clustered the 142 sample galaxies and then classified the mean spectrum in each cluster and reported how well that did.  But for the SOM, the authors constructed the SOM based on the 12 templates from K96 and then assigned classes to individual new spectra (not the nodes).  This doesn't seem comparable.  Please use the same methodology in both cases.
--- This is a reasonable criticism and we have ...

Further, k-means and the SOM used different distance functions (Euclidean vs. spectral angle distance) when operating on the same data.  Differences in results could simply be due to this change.  The authors' explanation for this choice does not make sense to me.

%TBD: --- We have results from the Euclidean distance for the K-means! But I steel think because of the way SOM operates it is better to use SAD! Or maybe describe it better!

The text says that Figure 11 shows "the best and the worst fitting results among the 22 clusters" (for k-means) but instead it seems to show the best-fit for two clusters (9 and 22) which both look pretty good.  The text implies that Fig 11 (right) is a poor match, but it looks very good.  Please explain.
--- We apologize for the confusion: this was a typo with "right" and "left" switched.

4. If k-means is retained, the way it is presented needs improvement. The description of k-means clustering in section 1 is a nice addition, but the comparison between k-means and SOM is confusing and wordy. K-means optimizes a partition of the data based on distance in the feature space only.  SOMs operate on a combination of distance in feature space and distance in the resulting low-dimensional map.  I am not sure how this achieves "smoothing", nor whether this is an important detail (the paper never mentions it again).  I recommend explaining what is meant here and why a "smooth" solution is better, or omitting this detail.  Explaining how SOM can be reduced to k-means is a good addition.
--- We have rewritten this paragraph (third to last of Section 1) and omitted the discussion of smoothing.

Please a add citation to the original k-means algorithm (e.g., MacQueen, 1967) when it is first mentioned (in addition to citing people who've used the algorithm).
--- Added the proper citation.

@@

Section 3.3 outlines the k-means method that is used in the experiments.  The language here appears to be more informal and less polished than that which describes the SOM method.  To explain k-means with the same level of detail, it would be appropriate to show the objective function that is employed and to explain that it uses an alternating optimization to assign points to centroids and then update the centroids, iterating until... what?  (There are different termination conditions used by different implementations, so it would be useful to indicate here what the matlab method does.)

--- We have reorganized Section 3 to discuss K-means first, followed by SOM. The former section 3.3 (now 3.1) has been rewritten, giving the objective function and using more formal language.

### 2017-05-21 main.tex
Pualine'e edits on version 8

\title[Classifying high-$z$ galaxy spectra]{Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach}
%\author{rahmani.sahar}
\date{\today}
\author[S.~Rahmani, H.~Teimoorinia and P.~Barmby]{S.~Rahmani$^{1,2}$\thanks{E-mail:
srahma49@uwo.ca}, H.~Teimoorinia$^{3}$, P.~Barmby$^{1,2}$\\
$^{1}$Department of Physics $\&$ Astronomy, Western University, London, ON N6A 3K7, Canada\\
$^{2}$Center for Planetary Science \& Exploration, Western University, London, ON N6A 3K7, Canada\\
$^{3}$Department of Physics $\&$ Astronomy, University of Victoria, Finnerty Road, Victoria, British Columbia, V8P 1A1, Canada}
\maketitle

@@

    The properties of the galaxies in the \citetalias{Hossein12} sample using the new classification were found to be more tightly correlated in mean values of age, specific star formation rate, stellar mass, and far-UV extinction than in previous studies. 
    The properties of the galaxies in each group are in good agreement with their morphological types.
    This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates; some of the spectra fall in between template classes. 
    This implies that these galaxies are not markedly different from those found in the local universe.
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/SFR and SSFR/M* relation.
     Other classification methods such as K-means and supervised artificial neural networks can provide a general classification with similar performance to the SOM. The ordered results and intermediate classes produced by self-organizing maps are distinct advantages of this method which can be utilized for other astronomical applications.

>>>>

\title[Classifying high-$z$ galaxy spectra]{Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach}
%\author{rahmani.sahar}
\date{\today}
\author[S.~Rahmani, H.~Teimoorinia and P.~Barmby]{S.~Rahmani$^{1,2}$\thanks{E-mail:
srahma49@uwo.ca}, H.~Teimoorinia$^{3}$, P.~Barmby$^{1,2}$\\
$^{1}$Department of Physics $\&$ Astronomy, Western University, London, ON N6A 3K7, Canada\\
$^{2}$Centre for Planetary Science \& Exploration, Western University, London, ON N6A 3K7, Canada\\
$^{3}$Department of Physics $\&$ Astronomy, University of Victoria, Finnerty Road, Victoria, British Columbia, V8P 1A1, Canada}
\maketitle

@@

    The properties of the galaxies in the \citetalias{Hossein12} sample using the new classification were found to be more tightly correlated in mean values of age, specific star formation rate, stellar mass, and far-UV extinction than in previous studies. 
    The properties of the galaxies in each group are in good agreement with their morphological types.
    This particular sample of high-redshift galaxies is well-described by the range of the \citetalias{Kinney96} templates; some of the spectra fall in between template classes. 
    This implies that these galaxies are not markedly different from those found in the local universe.
    The self-organizing map method can order the spectra in such a way as to also match the order along the age/SFR and SSFR/M* relation.
     Other classification methods such as K-means and supervised artificial neural networks can provide a general classification with similar performance to the SOM. The ordered results and intermediate classes produced by self-organizing maps are distinct advantages of this method which can be utilized for other astronomical applications.
%TBD: add some text here explainng why ordering is important

### 2017-05-21 sections0.08/intro0.08.tex
Pualine'e edits on version 8

%K-means vs. SOM
% the comparison between k-means and SOM is confusing and wordy. K-means optimizes a partition of the data based on distance in the feature space only.  SOMs operate on a combination of distance in feature space and distance in the resulting low-dimensional map.  I am not sure how this achieves "smoothing", nor whether this is an important detail (the paper never mentions it again).  I recommend explaining what is meant here and why a "smooth" solution is better, or omitting this detail.  Explaining how SOM can be reduced to k-means is a good addition.
K-means clustering~\citep{Macqueen67} is another unsupervised method used in many astronomical studies~\citep[e.g.][]{DAbrusco12,Ordov14,Boersma14,Aycha16}.
K-means clustering divides data into $K$ clusters, in such a way that each data point belongs to the cluster with the nearest mean value.
To first order, SOM and K-means work similarly: both algorithms operate on the distances between \boldit{N} $\in \Re^m$ objects defined by the properties under consideration.
The SOM algorithm also considers distances between objects in the low-dimensional map, as  controlled by a neighborhood function. 
When the radius of the neighbourhood function goes to zero, the SOM algorithm loses its ordering power and acts like K-means clustering.
%%%%After second referee report
\textbf{On the other note, in the K-means clustering the centre of each cluster are independent, but in the SOM the winner nodes adjust each other and provide an ordered display of clusters. Therefore, the (dis)-similarity between each cluster becomes more profound.}

%20170122: combined and reorganized last 2 paragraphs of section - DRAFT
%%%%After second referee report
The goal of this work is to compare SOM-based classification of galaxy spectra to two other methods: supervised neural network and (unsupervised) K-means. 
The SOM-based classification involves generating one- and two-dimensional self-organizing maps using the \citetalias{Kinney96} spectral templates. 
These maps are used to classify the spectra of 142 galaxies with $0.5 < z < 1$, the same
sample as analyzed by ~\citet[][hereafter \citetalias{Hossein12}]{Hossein12}.
That work classified spectra using a supervised (multi-layer feed-forward) neural network method and the spectral templates presented by \citetalias{Kinney96}.
The K-means method is used to cluster the \citetalias{Kinney96} and then find distance of each \citetalias{Hossein12} data to centre of clusters to classify them.
 As in \citetalias{Hossein12}, we examine the averaged properties of the new galaxy groups and compare them with previous work.
 %HERE IS WHERE WE MENTION OUR METRIC TO SAY WHICH METHOD IS BETTER

% haven't figured out where to put this.
%\citetalias{Hossein12} also 
%showed that there are tight correlations between physical properties of galaxies, and these %correlations might be different for each type of galaxy.
%This method could only match 105 out of the 142 galaxies directly to the \citetalias{Kinney96} templates; classifying the remaining 37 galaxies required combining template spectra.
%Our hypothesis was that since a self-organizing map has the freedom to classify objects in between known classes, it could be applied to these data to find the best spectral class for the 37 galaxies that could not be classified by the supervised method.

%What we did

>>>>

K-means clustering~\citep{Macqueen67} is another unsupervised method used in many astronomical studies~\citep[e.g.][]{DAbrusco12,Ordov14,Boersma14,Aycha16}.
K-means clustering divides data into $K$ clusters, in such a way that each data point belongs to the cluster with the nearest mean value.
To first order, SOM and K-means work similarly: both algorithms operate on the distances between \boldit{N} $\in \Re^m$ objects defined by the properties under consideration.
The SOM algorithm also considers distances between objects in the low-dimensional map, as  controlled by a neighborhood function. 
When the radius of the neighbourhood function goes to zero, the SOM algorithm loses its ordering power and acts like K-means clustering.

%20170122: combined and reorganized last 2 paragraphs of section - DRAFT
%%%%After second referee report
The goal of this work is to compare SOM-based classification of galaxy spectra to two other methods: supervised neural network and (unsupervised) K-means clustering.
The questions we want to answer are: (1) do galaxy spectra form a continuous distribution or can they be separated into discrete groups as represented by the \citetalias{Kinney96} templates? (2) if the distribution is continuous, how do galaxy properties vary along the spectral distribution.
We analyze the same sample of 142 galaxies with $0.5 < z < 1$ as 
\citet[][hereafter \citetalias{Hossein12}]{Hossein12}, who classified the spectra using a supervised (multi-layer feed-forward) neural network method with the \citetalias{Kinney96} templates.
This method could only match 105 out of the 142 galaxies directly to the \citetalias{Kinney96} templates; classifying the remaining 37 galaxies required combining template spectra.
%not sure if we want this sentence or not
%Our hypothesis was that since a self-organizing map has the freedom to classify objects in between known classes, it could be applied to these data to find the best spectral class for the 37 galaxies that could not be classified by the supervised method.
% also not sure if we want these sentences.
% As in \citetalias{Hossein12}, we examine the averaged properties of the new galaxy groups and compare them with previous work.
%\citetalias{Hossein12} also 
%showed that there are tight correlations between physical properties of galaxies, and these %correlations might be different for each type of galaxy.
 %HERE IS WHERE WE MENTION OUR METRIC TO SAY WHICH METHOD IS BETTER

### 2017-05-21 sections0.08/method0.08.tex
Pualine'e edits on version 8

\section{METHOD}
\label{sec: method_somz}
 \subsection{Self Organizing Maps}
 \label{sec: som}

@@

 \subsubsection{Algorithm of SOM} 
 \label{sec: algorithm}
     Assume we have a dataset which contains vectors, \boldit{V} $\in \Re^n$, and we want to map them on an S1 by S2 map. 
     Sizes of SOMs are arbitrary and there are no rules regarding choosing one over the other. 
    \citet{Vesanto05} suggested that a total number of $5\sqrt{n}$ neurons is a sufficient size, but users usually choose the size of the grids based on their dataset and their application of the results.

     We start by creating S1 $\times$ S2 empty neurons. 
     The arrangement of these neurons depends on the map's topology provided by the user. 
     In the case of 1D maps, since each neuron has two immediate neighbours, the topology of the map does not have any effect on the final result and any topology can be chosen.
     However, in 2D maps, the shape of the neurons specifies the number of immediate neighbours for each neuron and it is up to user to choose the most suitable shape based on the data.
     In this paper, we choose hexagonal topology, which gives each neuron six neighbours, and provides more interactions between neurons.
     Initially a random weight vector, \boldit{W} $\in \Re^n$, will be assigned to each node.
     The process of creating SOM happens over a series of $N$ iterations, where $N$ is set by the user. 
     During each iteration the weight vectors might change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
      In each iteration, the SOM code:
     \begin{enumerate}
        \item chooses a random vector from the dataset ($V_i$).
        \item calculates the Euclidean distance in $\in \Re^n$ space for each node, $j$, as  $D_j^2= \sum_{i=0}^{i=n} (V_i - W_{j_i})^2$, and finds a neuron with minimum $D_j$, (``$D_{j_{min}}$"). This neuron is the winner node and is called the Best Matching Unit (BMU). 
        \item  computes the radius of the neighbourhood of the BMU to find nodes within this radius. The weight vectors of these nodes will be affected in the next steps. The radius of the neighbourhood is arbitrary and can be set to be as high as half of the SOM size. It then decays exponentially over each iteration as
        \begin{equation}
            r^t_{BMU} = r^0_{BMU}e^{(-t/\tau)}
        \end{equation}
        where $\tau$ is a decay constant and is usually set to be the same as the number of iterations, $N$. $r^0_{BMU}$ and $r^t_{BMU}$ are the radii of the neighbourhood at 0th and $t$th iteration, respectively. 
        \item changes the weight vectors of the BMU and all the nodes within $r^t_{BMU}$ as:
        \begin{equation}
            \label{equ: weight adj}
            w(t+1)=w(t)+L(t) \times R(t) \times(v(t)-w(t))
        \end{equation}
        where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor, which prevents the divergence of the SOM and $R(t)=\exp(-\frac{D_j^2}{2r^t_{BMU}})$ is the influence rate. $R(t)$ determines how the weight of each node in the neighbourhood of BMU will change.
     \end{enumerate}
     These steps are then repeated $N$ times.

@@

     %%%After_second_referee_report
     \textbf{As mentioned in the introduction, one of the main advantages of the SOM method  is that the resulting networks can be used to cluster new datasets with no additional training required.
     New data with the same dimensionality as the original input data can be presented to the already trained network: the SOM algorithm finds the best matching unit considering the weight of nodes in the trained network and each vector from the new data set.
     The winner node determines the place of each new vector on the SOM.
     Based on the location of the new vectors on the SOM, they can be compared with the original data set. %20170122PB: unclear what "they" refers to
     Sections~\ref{sec: 1Dv} and ~\ref{sec: 2D} show the results of galaxy spectral classifications using networks trained on template spectra.}

\subsection{K-means clustering}
\label{sec: kmeans_method}     
In the K-means method, the user decides in advance on a number of clusters, $K$, which cannot be greater than the number of data points. 
The algorithm randomly chooses $K$ random points to be the centroids of the clusters.
Then it finds the data points that are closest to each assumed centroid and clusters them together.
The mean values of the data points assigned to each cluster are calculated and become the new centroid of the cluster. 
The algorithm finds new clusters based on the new centroids and repeats the procedure until it converges.
We used \textsc{matlab} K-means library~\citep{Seber84, Spath85}, written for \textsc{matlab2015b}, to perform the K-means clustering.

>>>>

\section{METHOD}
\label{sec: method_somz}

\textbf{In this section we describe the algorithms for both self-organizing maps and K-means clustering. As K-means is both simpler and more well-known, we introduce it briefly first before discussing SOM in detail.}

%Sahar: note that I changed the notation a little bit here, for both K-means and SOM. Where you had $n$ in the SOM description I changed it to $d$.

\subsection{K-means clustering}
\label{sec: kmeans_method}

\textbf{
The K-means method partitions a data set $x_i$, having $n$ points in $d$-dimensional feature space, into $K$ clusters.
The number of clusters $K<n$ must be decided by the user in advance, and the cluster centroids in feature space are located at positions $\mu_k$.
The goal of the algorithm is to minimize the sum
\begin{equation}
\label{eq:km_sos}
J = \sum_{i=1}^n \sum_{k=1}^K \min \big( \norm{x_i - \mu_k}^2 \big)
\end{equation}
where $\norm{z}$ indicates the distance measurement in $d$-dimensional space. In this work the Euclidean distance is is used, but other metrics are also possible.
The algorithm is initialized by randomly choosing $K$ random points as cluster centroids.
Each data point is assigned to the cluster with the closest centroid, and the centroids are re-calculated.
The steps of re-assigning cluster membership and re-calculating the cluster centroids are repeated until a stopping criterion is reached (a maximum number of iterations or no change in cluster assignments).
We used the \textsc{matlab} K-means library~\citep{Seber84, Spath85}, written for \textsc{matlab2015b}, to perform  K-means clustering.
The default maximum number of iterations (100) was used.
}

%Sahar: Please check that the last sentence is correct. Also please check that the K-means version you used was randomly picking the initial centroids - looks like Matlab can also do this in other ways.

 \subsection{Self Organizing Maps}
 \label{sec: som}

@@

 \subsubsection{Algorithm of SOM} 
 \label{sec: algorithm}
     Assume we have a dataset which contains vectors, \boldit{V} $\in \Re^d$, and we want to map them on an S1 by S2 map. 
     Sizes of SOMs are arbitrary and there are no rules regarding choosing one over the other. 
    \citet{Vesanto05} suggested that a total number of $5\sqrt{d}$ neurons is a sufficient size, but users usually choose the size of the grids based on their dataset and their application of the results.

     We start by creating S1 $\times$ S2 empty neurons. 
     The arrangement of these neurons depends on the map's topology provided by the user. 
     In the case of 1D maps, since each neuron has two immediate neighbours, the topology of the map does not have any effect on the final result and any topology can be chosen.
     However, in 2D maps, the shape of the neurons specifies the number of immediate neighbours for each neuron and it is up to user to choose the most suitable shape based on the data.
     In this paper, we choose hexagonal topology, which gives each neuron six neighbours, and provides more interactions between neurons.
     Initially a random weight vector, \boldit{W} $\in \Re^d$, will be assigned to each node.
     The process of creating SOM happens over a series of $N$ iterations, where $N$ is set by the user. 
     During each iteration the weight vectors might change according to the Kohonen learning rule (equation~\ref{equ: weight adj}). 
      In each iteration, the SOM code:
     \begin{enumerate}
        \item chooses a random vector from the dataset ($V_i$).
        \item calculates the Euclidean distance in $\in \Re^d$ space for each node, $j$, as  $D_j^2= \sum_{i=0}^{i=d} (V_i - W_{j_i})^2$, and finds a neuron with minimum $D_j$, (``$D_{j_{\min}}$"). This neuron is the winner node and is called the Best Matching Unit (BMU). 
        \item  computes the radius of the neighbourhood of the BMU to find nodes within this radius. The weight vectors of these nodes will be affected in the next steps. The radius of the neighbourhood is arbitrary and can be set to be as high as half of the SOM size. It then decays exponentially over each iteration as
        \begin{equation}
            r^t_{\rm BMU} = r^0_{\rm BMU}e^{(-t/\tau)}
        \end{equation}
        where $\tau$ is a decay constant and is usually set to be the same as the number of iterations, $N$. $r^0_{BMU}$ and $r^t_{\rm BMU}$ are the radii of the neighbourhood at 0th and $t$th iteration, respectively. 
        \item changes the weight vectors of the BMU and all the nodes within $r^t_{\rm BMU}$ as:
        \begin{equation}
            \label{equ: weight adj}
            w(t+1)=w(t)+L(t) \times R(t) \times(v(t)-w(t))
        \end{equation}
        where $L(t) = L_0 e^{(-t/\tau)}$ is the learning factor, which prevents the divergence of the SOM and $R(t)=\exp(-\frac{D_j^2}{2r^t_{\rm BMU}})$ is the influence rate. $R(t)$ determines how the weight of each node in the neighbourhood of BMU will change.
     \end{enumerate}
     These steps are then repeated $N$ times.

@@

     %%%After_second_referee_report
     \textbf{As mentioned in the introduction, one of the main advantages of the SOM method  is that the resulting networks can be used to cluster new datasets with no additional training required.
     New data with the same dimensionality as the original input data can be presented to the already trained network: the SOM algorithm finds the best matching unit considering the weight of nodes in the trained network and each vector from the new data set.
     The winner node determines the place of each new vector on the SOM.
     The location of the new vectors on the SOM allows them to be compared with the original data set. 
     Sections~\ref{sec: 1Dv} and ~\ref{sec: 2D} show the results of galaxy spectral classifications using networks trained on template spectra.}

### 2017-05-21 sections0.08/results0.08.tex
Pualine'e edits on version 8

       In order to check whether our clustering in Sec.~\ref{sec: 1Dv}~is correct, in this section we examine the relations between median properties of the galaxies in each neuron.

@@

       One of the key features of SOMs is the ability to not only classify objects but to provide an {\em ordered} classification. 
       Figure~\ref{fig: props_vs_props} shows that the SOM correctly determines an ordering of the physical properties of the galaxy models used to make the classified spectra. 
       The median properties of the spectral groups change on going from one end of the SOM to the other.
       K-means clustering and supervised ANNs can sort galaxy spectra into classes but do not naturally order them.
       Therefore, by using the SOM method, we not only study galaxies in different clusters, we also can study the trends between them as well.

>>>>

       In order to check whether our clustering in Sec.~\ref{sec: 1Dv}~is meaningful, in this section we examine the relations between median properties of the galaxies in each neuron.

@@

    % This is better but could perhaps still use work    
       \textbf{One of the key features of SOMs is the ability to not only classify objects but to provide an {\em ordered} classification.  K-means clustering and supervised ANNs can sort galaxy spectra into classes but do not naturally order them. Ordering is important in the case where the physical causes of the differences between classes/clusters are unknown. Average properties of objects within clusters (possibly from data not used for classification) can be used to test hypotheses about the physical differences. 
       Figure~\ref{fig: props_vs_props} shows that the SOM ordering also corresponds to the ordering of the physical properties of the galaxy models used to make the classified spectra. 
       The median properties of the spectral groups change on going from one end of the SOM to the other.
       Therefore, by using the SOM method, we can study not only how galaxies group into spectral clusters, but the trends between the clusters as well.}
